\section{Motivation}
\label{sec:motivation}

For concurrency bugs, multiple code fragments can be involved, possibly in
different units of code. In this ambiguous scenario the appropriate fix is not
always clear. There has been work in concurrency anti-patterns that provide a
definition, problem, context and solution for concurrency
bugs~\cite{BJ09,FKLV12}.

We now provide a simple example of a data race bug and how a developer might
fix it. In the left part of Fig.~\ref{fig:fixed_sample_datarace} the
\texttt{read} and \texttt{write} method access a shared variable. A very simple
data race exists because there is no atomic access to the \texttt{data}
variable during the concurrent reading or writing.
%Both methods are involved in the data race, and it is because of the interactions between these methods the data race is possible.
A solution involves synchronizing both accesses as shown in the right part of
Fig.~\ref{fig:fixed_sample_datarace}. Synchronizing one method alone does not
completely fix the bug in this case.

% Phase 2 - the optimization phase in development - addresses the issue of unnecessarily
% synchronizing the ``expensive loop'' and ``database query''.

The solution in the right part of Fig.~\ref{fig:fixed_sample_datarace} is far
from ideal. It forces other threads to wait unnecessarily while the write
method works in the loop and database sections. An optimization is to shrink
the critical region (the synchronized statements) to only guard access to the
shared variable as shown in Fig.~\ref{fig:optimized_sample_datarace}.

\begin{figure}[h]
\begin{minipage}{5cm}
\footnotesize{\textbf{Buggy Program:}}
\begin{lstlisting}[language=Java, morekeywords={synchronize}]
write(int var1){
  ... // Expensive loop
  data = var1;
  ... // Database query
}

int public read(){
  return data;
}
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{5cm}
\footnotesize{\textbf{Fixed Program:}}
\begin{lstlisting}[language=Java, morekeywords={synchronize}]
synchronize write(int var1){
  ... // Expensive loop
  data = var1;
  ... // Database query
}

int synchronize read(){
  return data;
}
\end{lstlisting}
\end{minipage}
\caption{A developer first synchronizes the \texttt{read} function, yet the bug
still exists. Synchronizing the \texttt{write} method as well fixes it.}
\label{fig:fixed_sample_datarace}
\end{figure}

\begin{figure}[t]
\begin{minipage}{5cm}
\footnotesize{\textbf{$1^{st}$ Optimization on Fix:}}
\begin{lstlisting}[language=Java, morekeywords={synchronize}]
public write(int var1){
 ... // Expensive loop
 synchronized(this){
   data = var1;
   ... // Database query
 }
}

int synchronize read(){
 return data;
}
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{5cm}
\footnotesize{\textbf{$2^{nd}$ Optimization on Fix:}}
\begin{lstlisting}[language=Java, morekeywords={synchronize}]
public write(int var1){
 ... // Expensive loop
 synchronized(this){
   data = var1;
 }
 ... // Database query
}

int synchronize read(){
 return data;
}
\end{lstlisting}
\end{minipage}
\caption{A developer shrinks the critical region to exclude the expensive loop.
They shrink the critical region again to exclude the database query as a
further optimization.}
\label{fig:optimized_sample_datarace}
\end{figure}

From a developer's standpoint there is a lot of work involved in creating a fix
for parallel bugs as multiple, unrelated changes are not uncommon. Two changes
were required to functionally fix the example program, and two additional
changes were required to improve the non-functional performance. An ideal fix
only requires two changes which minimizes the time spent in the critical
region. ARC currently finds a fix for a concurrency bug, it does not make any
attempt at ensuring it is the optimized fix (this is ongoing research explained
in Sect.~\ref{sec:ongoing}).

Automated tools in software testing and debugging are needed as they have the
potential to reduce the vast amount of resources spent on software testing
(upwards to \$59.5 billion)~\cite{RTI02}.
%ARC provides an automated approach to fixing the \textit{functionality}
%, and then optimizing the \textit{non-functional} performance
%of programs with deadlocks and data races.

\begin{quote}
\textit{Some properties are difficult or impossible to encode using test cases,
such as nondeterministic properties; GenProg cannot currently repair race
conditions, for example.}~\cite{GNFW11}
\end{quote}
